---
title: 'Reproducible Research: Peer Assessment 1'
author: "Renato Deggau - rdeggau@gmail.com"
date: "Tuesday, August 11, 2015"
output: html_document
keep_md: true
---
## Loading and preprocessing the data
```{r loading, echo = TRUE}
monitoring <- read.csv("./activity.csv", header=TRUE)
summary(monitoring)
```

## What is mean total number of steps taken per day?
### histogram and mean and median

```{r histogram, echo = TRUE}
library(ggplot2)
monitoringHist <- aggregate(x = monitoring$steps , by = list(monitoring$date), FUN = sum ,na.rm=TRUE)
names(monitoringHist) <- c("date","steps")
histplot <- ggplot(monitoringHist,aes(x = steps)) +
            ggtitle("Histogram of Number of Steps By Day") +
            xlab("Steps") +
            geom_histogram(binwidth = 2500)
histplot
mean(monitoringHist$steps)
median(monitoringHist$steps)
```

## What is the average daily activity pattern?
```{r averageDaily, echo = TRUE}
activity <- aggregate(x = monitoring$steps , by = list(monitoring$interval), 
                      FUN = mean , na.rm=TRUE)
names(activity) <- c("interval","steps")
lineplot <-  ggplot(activity, aes(interval,steps)) +
                 ggtitle("Average Steps - Time Series") +
                 geom_line()
lineplot
activity$interval[which.max(activity$steps)]
```

## Imputing missing values
### number of missing values
```{r missingValues, echo = TRUE}
sum(is.na(monitoring))
```
### fill missing values - use the mean
```{r fillmissingValues, echo = TRUE}
monitoringClean <- monitoring
monitoringClean$steps[is.na(monitoringClean$steps)] <- 
    tapply(monitoring$steps, monitoring$interval, mean, na.rm = TRUE)
summary(monitoringClean)
sum(is.na(monitoringClean))
```
### new histogram with new data
```{r newhistogram, echo = TRUE}
library(ggplot2)
monitoringCleanHist <- aggregate(x = monitoringClean$steps , 
                         by = list(monitoringClean$date), FUN = sum ,na.rm=TRUE)
names(monitoringCleanHist) <- c("date","steps")
histplot <- ggplot(monitoringCleanHist,aes(x = steps)) +
            ggtitle("Histogram of Number of Steps By Day Without Missing Values") +
            xlab("Steps") +
            geom_histogram(binwidth = 2500)
histplot
mean(monitoringCleanHist$steps)
median(monitoringCleanHist$steps)
```

## Are there differences in activity patterns between weekdays and weekends?
### Create weekdays and weekends variable
```{r weekdays, echo = TRUE}
daytype <- function(date) {
    if (weekdays(as.Date(date)) %in% c("Saturday", "Sunday")) {
        "weekends"
    } else {
        "weekdays"
    }
}
monitoringClean$daytype <- as.factor(sapply(monitoringClean$date, daytype))

par(mfrow = c(2, 1))
for (type in c("weekends", "weekdays")) {
    monitoringWeek <- aggregate(x = monitoringClean$steps , 
                                by = list(monitoringClean$interval), 
                                subset = monitoringClean$daytype == type,
                                FUN = mean , na.rm=TRUE) 
    plot(monitoringWeek, type = "l", main = type)
}
 
```
